Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not easily perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard). We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand the spatial regions humans refer to. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.

This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

@@

The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e. what can be done there and where. In this work we implicitly represent context using the types of objects present in a room and their location, both relative to each other and relative to the room itself. The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.

@@

Dora is capable of using vision to recognize pre-trained 3D object models. Recognition can either be triggered through autonomous visual search or at a user's command. When an object is detected it is represented in the metric map 
by placing a copy of the model at the detected pose. The recognizer associates each object with a type label that was provided during a training phase.

@@

We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs. Anchor points are symbolic descriptions which link a conceptual entity (i.e. something that can be reasoned about but not directly perceived, such as a CDSR) to a particular point on a perceived entity. The perceived entities we use are the objects recognised by Dora, and the room itself. The room representation is created by putting a convex hull around the lines in Dora's SLAM map. Anchor points are created from perceived entities using unary functions, e.g. \fw{(CentroidFn Desk1)} represents the centroid of the geometric extent of \fw{Desk1}, and are linked to a particular CDSR using a \fw{regionBoundary} relation. After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation. Therefore, each CDSR has one type and a variable number of boundary points.

\from{Nick}{TODO: Need to update example, including new fns, when we have updated pictures in.}

@@

\fw{(regionType CDSR1 FrontRegion)\\
(regionBoundary CDSR1 (LeftmostTopFn Room1))\\
(regionBoundary CDSR1 (LeftmostBottomFn Room1))\\
(regionBoundary CDSR1 Desk1)\\
(regionBoundary CDSR1 Desk2)\\
(regionBoundary CDSR1 Desk3)\\
(regionBoundary CDSR1 Desk4)\\
(regionBoundary CDSR1 Whiteboard0)
}}
  \caption{Representing the front of the classroom context-dependent region \fw{CDSR1}}
  \label{fig:cdsr-reps}
\end{figure}

@@

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
  \caption{Analogical mapping between five base expressions and two target expressions. }
  %Ovals and squares represent relations and entities respectively. Black dashed bi-directional arrows are the correspondences between the base and target, and the red dashed objects represent a candidate inference projected into the target.

@@

\from{Nick}{Kate, please update figure caption, plus descriptive paragraph below}

Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and two target ones. Each oval represents a binary predicate, and its two entity arguments are represented by squares. SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)}. This is done in the following manner. First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME. Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11} and \fw{Desk2} with \fw{Desk12}. While there is another \fw{below} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa. In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows. Next, SME creates a candidate inference for the region boundary expression, because \fw{Desk1} participates in the mapping. The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}. \fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences. In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes the \fw{XMinYMost} point of the two desk group. Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions. %Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.

@@

% \section{Acknowledgments}
% 
% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams. The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.

>>>>

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard). We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.

\from{klenk}{Cite recent Langley work}

This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) corresponding to sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically and geometrically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

@@

The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e. what can be done there and where. In this work we implicitly represent context using the types of objects present in a room and their location relative to each other. The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.

@@

Dora is capable of using vision to recognize pre-trained 3D object models. Recognition can either be triggered through autonomous visual search or at a user's command. When an object is detected it is represented in the metric map 
by placing a copy of the model at the detected pose. The recognizer associates each object with a semantic type that was provided during a training phase.

@@

We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs. Anchor points are symbolic descriptions which link a conceptual entity to a perceived entity. The perceived entities we use are the objects recognised by Dora, and the room itself. The room representation is created by putting a convex hull around the lines in Dora's SLAM map. Anchor points are created from perceived entities using unary functions, e.g. \fw{(XMaxYMostFn Desk1)} represents the point on the \fw{Desk1} with the largest x coordinate taken from the set of points with a y coordinate within 5\% of the maximum y coordinate. Anchor points are linked to particular CDSRs using a \fw{regionBoundary} relation. After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation. Therefore, each CDSR has one type and a variable number of boundary segments.

@@

\fw{
(regionType CDSR9 FrontRegion) \\
(regionBoundarySegment CDSR9 \\
\hspace*{2em}(YMaxXFewestFn Room-Base) \\
\hspace*{2em}(YMinXFewestFn Room-Base)) \\
(regionBoundarySegment CDSR9 \\
\hspace*{2em}(YMinXFewestFn Room-Base) \\
\hspace*{2em}(YMinXFewestFn Group1))\\
...
}
}
  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
  \label{fig:cdsr-reps}
\end{figure}

@@

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{images/AnalogyDiagram.png}
  \caption{Analogical mapping between five base expressions and three target expressions. }
  %Ovals and squares represent relations and entities respectively. Black dashed bi-directional arrows are the correspondences between the base and target, and the red dashed objects represent a candidate inference projected into the target.

@@

Figure \ref{fig:analogy} illustrates a sample mapping between five base expressions and three target ones. Each oval represents a binary predicate, and its two entity arguments are represented by squares. SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case . This is done in the following manner. First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME. Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}. While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa. In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows. Next, SME creates a candidate inference for the region boundary expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping. The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}. \fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences. In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes the \fw{XMinYMost} point of the two desk group. Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions. %Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.

@@

% \section{Acknowledgments}
% 
% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams, Scott Friedman for assistance on probablistic SME, and Danny Bobrow comments on an earlier draft of this paper. The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.
